Gravitational wave sources in our Galactic backyard: LISA predictions for BH and NS binaries

### 2021-06-10 paper/tex/sections/discussion.tex
some of pulsar matching

\subsection{Matching LISA detection to radio pulsars}\label{sec:pulsar_matching}

>>>>

\subsection{Matching LISA detection to radio pulsars}\label{sec:pulsar_matching}
Since the vast majority of the population will not merge for many years, the main type of electromagnetic counterparts for the LISA Galactic DCO population is pulsars. Therefore, for this section we focus only on BHNSs and NSNSs since no BHBH system will contain a pulsar. In this section we investigate how we can use SKA to match pulsars to LISA signals with some back-of-the-envelope estimates.

First, we can consider how many pulsars SKA is likely to detect. \citet{Keane+2015} uses PSRPOPPy \citep{Bates+2014} to simulate the Milky Way pulsar population. They find that for SKA-1, approximately $10000$ pulsars will be discovered. The second phase of SKA should also have been approved and in operation by the time of the LISA mission and this phase would yield a total of $24000$-$30000$ pulsars \citep{Keane+2015}. Moreover, we are only interested in pulsars that are part of a binary system. We can estimate the pulsar binary fraction as the fraction of known pulsars that are in binaries using the ATNF Pulsar Catalogue\footnote{\url{https://www.atnf.csiro.au/research/pulsar/psrcat}} \citep{Manchester+2005}. $290$ of the $2872$ currently known pulsars are in binary systems and thus we can estimate the binary fraction of pulsars as $10\%$.

Next, we find the total number of pulsars SKA will detect in a patch on the sky. The total sky area that the SKA mission covers is approximately $7200 \unit{deg^2}$, which is calculated based on SKA-mid covering Galactic latitudes of $\abs{b} < 10^\circ$ and SKA-low covering $\abs{b} < 5^\circ$ \citep{Keane+2015}. If we assume that the pulsars are found uniformly across the sky, this means that roughly $0.14$ and $0.42$ binary pulsars are expected per square degree for SKA-1 and SKA-2 respectively. Note that this assumption is not realistic as pulsars will tend to be far more concentrated in the Galactic centre but we use it to provide an upper bound on these estimates.

Given these estimates, and by considering the last panel of Fig.~\ref{fig:fiducial_pdf_distributions}, approximately $10$ and $6$ (for SKA-1 and SKA-2) DCOs containing NSs will be localised well enough that \textit{if} the NS is a pulsar, SKA can unambiguously match it to the radio signal.

Furthermore, both SKA and LISA will measure the frequency, frequency derivative and chirp mass of each of these systems. Therefore, they can independently calculate these parameters and use them to match a GW signal to a radio signal. In this way, there can be more than one pulsar within the sky localisation provided to LISA and it is still possible to match to the correct pulsar.
